---
title: "lab3 - gapminder"
author: "Matthew Pruden"
date: "February 26, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Data

Today, we’ll use ggplot to visually explore global trends in public health and economics compiled by the Gapminder project. This project was pioneered by Hans Rosling, who is famous for describing the prosperity of nations over time through famines, wars and other historic events with this beautiful data visualization in his 2006 TED Talk: The best stats you’ve ever seen.

<br>

###Loading data
```{r, message=FALSE}
library(dslabs)
library(tidyverse)
```

###Exploring fulldata data
```{r}
head(gapminder)
```

###Subsetting data for 2011
```{r}
gap2011 <- gapminder %>% 
  as_tibble() %>% 
  filter(year == 2011)

head(gap2011)
```

<br>

##Life Expectancy

###Life expectancy as a function of income
```{r}
ggplot(data = gap2011, aes(x = gdp/population, y = life_expectancy)) +
  geom_point( aes(, size = log(population), fill = continent), shape = 21, stroke = 1, color = "black") +
  xlab("Per Capita GDP") + 
  ylab("Life Expectancy") 
```

<br>

###Life expectancy by continent
```{r}
ggplot(data = gap2011, aes(x = continent, y = life_expectancy)) + 
  geom_boxplot(aes(color=continent)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  ylab("Life Expectancy") + 
  xlab("") + 
  ggtitle("Life Expectancy by Continent")
```

<br>

###Life expectancy by region
```{r}
ggplot(data = gap2011, aes(x = reorder(region, life_expectancy, FUN = median), y = life_expectancy)) +
  geom_boxplot(aes(color=continent)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  ylab("Life Expectancy") + 
  xlab("") + 
  ggtitle("Life Expectancy by Region")
```


##Fertility
```{r}
ggplot(data = gap2011, aes(x = fertility)) + 
  geom_density(aes(group = continent, fill = continent), alpha = 0.5) +
  ggtitle("Female Fertility as a Function of Per Capita GDP") + 
  ylab("Fertility") + 
  xlab("")
```

<br>

##Infant Mortality
```{r}
ggplot(data = gap2011, aes(x = gdp/population, y = infant_mortality)) + 
  geom_point(aes(color=continent), size = 2) +
  ggtitle("Infant Mortality as a Function of Per Capita GDP") + 
  ylab("Infant Mortality") + 
  xlab("Per Capita GDP")
```


```{r}
ggplot(data = gap2011, aes(x = infant_mortality, y = life_expectancy)) + 
  geom_point(aes(color = continent), size = 2) +
  geom_smooth(method='lm') + 
  ggtitle("The Effect of Infant Mortality on Life Expectancy") +
  ylab("Life Expectancy") +
  xlab("Infant Mortality")
```



